---
title: "`r params$crop_name` planting date and cultivar recommendations — `r params$country_name`"
output:
  html_document:
    self_contained: true
    theme:
      bootswatch: flatly
      version: 5
    mathjax: null
params:
  data_rds: ""
  crop_name: ""
  country_name: ""
  has_crop_mask: FALSE
  project_root: "."
---

<style>
/* rmarkdown's default template caps content at 940px regardless of theme -
   override it so the filters/maps/table can use more of a wide monitor. */
.main-container { max-width: 1600px; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages({
  library(crosstalk)
  library(leaflet)
  library(plotly)
  library(DT)
  library(dplyr)
  library(ggplot2)
  library(htmltools)
})
```

```{r map-helpers, include=FALSE}
# Builds one independent leaflet widget bound once to `sd`. Deliberately
# exactly one crosstalk-bound addCircleMarkers layer per widget INSTANCE -
# leaflet's crosstalk filter handling (htmlwidgets/assets/leaflet.js) tracks
# marker visibility per (crosstalk group, row key) assuming a single layer
# per key; binding multiple layers to the same group with the same row_key
# values within ONE widget (as an earlier version of this file did, for a
# switchable single-map coloring) breaks its internal bookkeeping and
# silently stops filtering from working at all. Calling this function
# multiple times - once per map below - produces separate widgets, each
# with its own independent JS layer registry, so there's no collision.
#
# Marker colors use a FIXED scale (the full crop's data range for that
# metric) on every map, kept constant across filter changes so a given
# color always means the same thing no matter what's filtered - only each
# map's live text readout (built later, via the map-lookups/script blocks)
# reflects the current filter state, rather than re-coloring markers, since
# dynamically restyling already-rendered canvas markers would need custom
# JS reaching into leaflet's internal layer registry that can't be verified
# without a live browser.
#
# Two outline layers are drawn, styled distinctly: the country border (thin
# black) and, when this run covers only part of the country, the zone
# border (thicker blue) - so a colleague can see at a glance both where the
# study area sits within the country and its own precise extent.
build_dssat_map <- function(sd, element_id, color_formula, pal_fn, pal_values,
                             legend_title, label_formula, outline_parts,
                             zone_outline_parts = list(), bounds = NULL,
                             leg_lab_format = NULL) {
  m <- leaflet::leaflet(
      sd, elementId = element_id, height = 480,
      options = leaflet::leafletOptions(preferCanvas = TRUE)) %>%
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) %>%
    leaflet::addCircleMarkers(
      lng = ~LONG, lat = ~XLAT, radius = 5, stroke = FALSE, fillOpacity = 0.8,
      color = color_formula, label = label_formula)

  m <- if (is.null(leg_lab_format)) {
    leaflet::addLegend(m, "bottomright", pal = pal_fn, values = pal_values, title = legend_title)
  } else {
    leaflet::addLegend(m, "bottomright", pal = pal_fn, values = pal_values,
                        title = legend_title, labFormat = leg_lab_format)
  }

  for (part in outline_parts) {
    m <- leaflet::addPolylines(
      m, lng = part$long, lat = part$lat, color = "black", weight = 1, opacity = 0.6)
  }
  for (part in zone_outline_parts) {
    m <- leaflet::addPolylines(
      m, lng = part$long, lat = part$lat, color = "#2c7fb8", weight = 2.5, opacity = 0.9)
  }

  # Explicit initial view, fit to where the simulated points actually are -
  # without this, leaflet auto-fits to every added layer including the
  # (much larger) country/zone outline polylines, so the map opens zoomed
  # out to the whole country instead of the study area.
  if (!is.null(bounds)) {
    m <- leaflet::fitBounds(m, bounds$lng1, bounds$lat1, bounds$lng2, bounds$lat2)
  }

  m
}

# One small bordered column: a label, a map widget, and its live readout -
# keeps the three maps visually grouped together for the "check rank 1, see
# all three at once" workflow, with each map's label sitting right with it
# rather than as a page-level heading.
map_col <- function(label, map_widget, readout_id, readout_label) {
  htmltools::div(
    htmltools::tags$h5(label),
    map_widget,
    htmltools::div(
      style = "margin-top: 4px; font-weight: bold; font-size: 0.9em;",
      paste0(readout_label, ": "), htmltools::tags$span(id = readout_id, "computing…"))
  )
}

# Administrative boundary outline (country, level 0, or a named zone/admin-1
# unit), for context on the maps below. Prefers a boundary file already
# cached on disk (reused from elsewhere in this pipeline) over a fresh
# geodata::gadm() network fetch, so rendering an already-run usecase's
# dashboard doesn't depend on network access. `zones`, when given, restricts
# the result to admin-1 units whose name matches one of the requested zone
# labels (partial, case-insensitive match, since a usecase's zone label -
# e.g. "Central" - doesn't always exactly match GADM's own naming - e.g.
# "Central Region") and dissolves them into a single outer boundary rather
# than showing every internal sub-district line.
fetch_admin_outline <- function(country_name, project_root, level, zones = NULL) {
  tryCatch({
    country_code <- countrycode::countrycode(
      country_name, origin = "country.name", destination = "iso3c")
    admin_dir <- file.path(project_root, "data", "countries", country_code, "admin")
    cache_file <- file.path(
      admin_dir, "gadm", paste0("gadm41_", country_code, "_", level, "_pk.rds"))
    adm <- if (file.exists(cache_file)) {
      terra::vect(cache_file)
    } else {
      geodata::gadm(country = country_code, level = level, path = admin_dir)
    }
    if (!is.null(zones) && length(zones) > 0) {
      name_col <- paste0("NAME_", level)
      pattern <- paste(zones, collapse = "|")
      adm <- adm[grepl(pattern, adm[[name_col]][[1]], ignore.case = TRUE), ]
      if (nrow(adm) == 0) return(NULL)
      adm <- terra::aggregate(adm)
    }
    adm_sf <- sf::st_as_sf(adm)
    coords <- sf::st_coordinates(adm_sf)
    data.frame(
      long = coords[, "X"], lat = coords[, "Y"],
      group = paste(coords[, "L2"], coords[, "L1"], sep = "."))
  }, error = function(e) NULL)
}
```

```{r load-data}
dat <- readRDS(params$data_rds)

cultivar_order <- intersect(c("Short", "Medium", "Long", "Longer"), unique(as.character(dat$Cultivar)))
dat$Cultivar <- factor(dat$Cultivar, levels = cultivar_order)
dat$PDAT <- as.Date(dat$PDAT)

display_cols <- c(
  "zone", "Combination_Rank", "TRNO", "XLAT", "LONG", "Cultivar", "PDAT",
  "HWAH", "CWAM", "in_crop_mask", "row_key")
dat <- dat[, display_cols]

sd <- crosstalk::SharedData$new(dat, key = ~row_key, group = "dssat")
```

This dashboard presents planting-date and cultivar maturity-class
recommendations for **`r params$crop_name`** in **`r params$country_name`**,
generated using the DSSAT crop simulation model driven by seasonal forecast
weather data (not historical or observed weather). At each simulated
location, the model was run across 8 planting-date treatments — one week
apart, beginning at the estimated onset of the rainy season — combined with
`r length(cultivar_order)` cultivar maturity classes
(`r paste(cultivar_order, collapse = ", ")`), for up to
`r 8 * length(cultivar_order)` simulated planting-date/cultivar combinations
per location.

This file is self-contained and can be opened directly in a web browser,
without a server or R installation; only the background map imagery
requires an internet connection. If this analysis covered more than one
crop, each has its own dashboard file in its respective results folder.

`r if (!isTRUE(params$has_crop_mask)) paste0("**Note:** a cropland mask is not yet available for ", params$crop_name, " in this region. The \"Crop mask\" filter below therefore has no effect, and all simulated locations are shown.")`

## Filters

```{r filters}
crosstalk::bscols(
  widths = c(6, 6),
  crosstalk::filter_checkbox("mask_filter", "Crop mask (checked = location is inside the mask)", sd, ~in_crop_mask),
  crosstalk::filter_checkbox("cultivar_filter", "Cultivar", sd, ~Cultivar)
)
```

<script>
// Crop mask defaults to checked on load. crosstalk::filter_checkbox has no
// default-checked parameter, so this simulates a user click: waiting for
// `load` (not `DOMContentLoaded`) guarantees htmlwidgets has already
// attached its delegated change listener before the simulated click fires.
window.addEventListener('load', function() {
  var input = document.querySelector('#mask_filter input[value="TRUE"]');
  if (input && !input.checked) {
    input.checked = true;
    input.dispatchEvent(new Event('change', { bubbles: true }));
  }
});
</script>

**Recommendation rank** orders each location's planting-date/cultivar
combinations by simulated yield, with 1 as the best-performing option and
higher numbers indicating progressively lower-yielding alternatives. Select
**1** to view only each location's top recommendation, or leave all ranks
selected to see every simulated combination.

```{r rank-filter}
crosstalk::filter_checkbox("rank_filter", "Recommendation rank", sd, ~Combination_Rank, inline = TRUE)
```

## Maps

```{r maps}
label_formula <- ~sprintf("%s | %s | %.0f kg/ha", Cultivar, format(PDAT, "%Y-%m-%d"), HWAH)

country_outline_df <- fetch_admin_outline(params$country_name, params$project_root, level = 0)
if (is.null(country_outline_df)) {
  country_outline_df <- tryCatch(
    ggplot2::map_data("world", region = params$country_name),
    error = function(e) NULL)
}
outline_parts <- if (!is.null(country_outline_df) && nrow(country_outline_df) > 0) {
  split(country_outline_df, country_outline_df$group)
} else {
  list()
}

# If this run covers one or more named zones rather than the whole country,
# draw those zones' own boundary too (see fetch_admin_outline() above).
zones_present <- sort(unique(as.character(dat$zone)))
zone_outline_df <- if (length(zones_present) > 0) {
  fetch_admin_outline(params$country_name, params$project_root, level = 1, zones = zones_present)
} else {
  NULL
}
zone_outline_parts <- if (!is.null(zone_outline_df) && nrow(zone_outline_df) > 0) {
  split(zone_outline_df, zone_outline_df$group)
} else {
  list()
}

# Initial map view: fit to where the simulated locations actually are, not
# the (typically much larger) country/zone outline extent.
lng_range <- range(dat$LONG, na.rm = TRUE)
lat_range <- range(dat$XLAT, na.rm = TRUE)
lng_pad <- max(diff(lng_range) * 0.1, 0.05)
lat_pad <- max(diff(lat_range) * 0.1, 0.05)
map_bounds <- list(
  lng1 = lng_range[1] - lng_pad, lat1 = lat_range[1] - lat_pad,
  lng2 = lng_range[2] + lng_pad, lat2 = lat_range[2] + lat_pad)

pal_yield <- leaflet::colorNumeric("YlOrRd", domain = dat$HWAH, na.color = "transparent")
pal_cultivar <- leaflet::colorFactor("Dark2", domain = cultivar_order, na.color = "transparent")
pal_pdat <- leaflet::colorNumeric("RdYlBu", domain = as.numeric(dat$PDAT), na.color = "transparent", reverse = TRUE)
# Day-month, no year (e.g. "7 - Aug") - cuts are full days-since-epoch, so a
# season spanning a year boundary (e.g. Nov -> Feb) still positions/orders
# correctly on the color scale; only the displayed label drops the year.
pdat_leg_labFormat <- function(type, cuts, p) {
  d <- as.Date(cuts, origin = "1970-01-01")
  paste0(as.integer(format(d, "%d")), " - ", format(d, "%b"))
}

yield_map <- build_dssat_map(
  sd, "dssat_map_yield", ~pal_yield(HWAH), pal_yield, dat$HWAH,
  "Yield (kg/ha)<br>(fixed color scale)", label_formula, outline_parts,
  zone_outline_parts, map_bounds)

cultivar_map <- build_dssat_map(
  sd, "dssat_map_cultivar", ~pal_cultivar(Cultivar), pal_cultivar, cultivar_order,
  "Cultivar<br>(fixed color scale)", label_formula, outline_parts,
  zone_outline_parts, map_bounds)

pdat_map <- build_dssat_map(
  sd, "dssat_map_pdat", ~pal_pdat(as.numeric(PDAT)), pal_pdat, as.numeric(dat$PDAT),
  "Planting date<br>(fixed color scale)", label_formula, outline_parts,
  zone_outline_parts, map_bounds, leg_lab_format = pdat_leg_labFormat)

crosstalk::bscols(
  widths = c(4, 4, 4),
  map_col("Estimated Yields", yield_map, "yield-range-text", "Visible yield range"),
  map_col("Recommended Cultivar", cultivar_map, "cultivar-counts-text", "Visible cultivar counts"),
  map_col("Recommended Planting Date", pdat_map, "pdat-range-text", "Visible planting-date range")
)
```

```{r map-lookups, include=FALSE}
# key -> value lookups consumed by the dynamic-legend script below. Kept out
# of the visible chunks (include=FALSE) since they're only for the JS, not
# for display.
yield_by_key_json <- jsonlite::toJSON(setNames(as.list(dat$HWAH), dat$row_key), auto_unbox = TRUE)
cultivar_by_key_json <- jsonlite::toJSON(setNames(as.list(as.character(dat$Cultivar)), dat$row_key), auto_unbox = TRUE)
# Plain ISO YYYY-MM-DD strings, not JS Date objects - they sort correctly
# with ordinary string min/max, so no date parsing is needed client-side.
pdat_by_key_json <- jsonlite::toJSON(setNames(as.list(format(dat$PDAT, "%Y-%m-%d")), dat$row_key), auto_unbox = TRUE)
cultivar_levels_json <- jsonlite::toJSON(cultivar_order)
```

<script>
window.dssatYieldByKey = `r yield_by_key_json`;
window.dssatCultivarByKey = `r cultivar_by_key_json`;
window.dssatPdatByKey = `r pdat_by_key_json`;
window.dssatCultivarLevels = `r cultivar_levels_json`;

(function() {
  function fmt(x) { return Math.round(x).toLocaleString(); }

  function valuesFor(lookup, keys) {
    if (keys === null || typeof keys === "undefined") {
      return Object.values(lookup);
    }
    return keys
      .map(function(k) { return lookup[k]; })
      .filter(function(v) { return typeof v !== "undefined"; });
  }

  function updateYield(keys) {
    var el = document.getElementById("yield-range-text");
    if (!el) return;
    var vals = valuesFor(window.dssatYieldByKey, keys).filter(function(v) { return typeof v === "number" && !isNaN(v); });
    if (vals.length === 0) { el.textContent = "no locations match the current filters"; return; }
    var lo = Math.min.apply(null, vals);
    var hi = Math.max.apply(null, vals);
    el.textContent = fmt(lo) + " – " + fmt(hi) + " kg/ha (" + vals.length + " points)";
  }

  function updatePdat(keys) {
    var el = document.getElementById("pdat-range-text");
    if (!el) return;
    var vals = valuesFor(window.dssatPdatByKey, keys).filter(function(v) { return typeof v === "string"; });
    if (vals.length === 0) { el.textContent = "no locations match the current filters"; return; }
    vals.sort();
    el.textContent = vals[0] + " – " + vals[vals.length - 1] + " (" + vals.length + " points)";
  }

  function updateCultivar(keys) {
    var el = document.getElementById("cultivar-counts-text");
    if (!el) return;
    var vals = valuesFor(window.dssatCultivarByKey, keys).filter(function(v) { return typeof v === "string"; });
    var counts = {};
    window.dssatCultivarLevels.forEach(function(lvl) { counts[lvl] = 0; });
    vals.forEach(function(v) { if (counts.hasOwnProperty(v)) counts[v] += 1; });
    el.textContent = window.dssatCultivarLevels
      .map(function(lvl) { return lvl + ": " + counts[lvl]; })
      .join(" · ");
  }

  function updateAll(keys) {
    updateYield(keys);
    updatePdat(keys);
    updateCultivar(keys);
  }

  // Mirrors the exact crosstalk.FilterHandle pattern leaflet's own bundled
  // JS uses internally (new crosstalk.FilterHandle(group), .on("change", ...),
  // reading .filteredKeys for the initial state) - reused here directly
  // rather than reverse-engineered from scratch, and confirmed (by reading
  // crosstalk.js's EventEmitter-based implementation) that a single
  // FilterHandle can safely drive multiple readouts from one subscription.
  function init() {
    if (typeof crosstalk === "undefined") { setTimeout(init, 200); return; }
    var filterHandle = new crosstalk.FilterHandle("dssat");
    filterHandle.on("change", function(e) { updateAll(e.value); });
    updateAll(filterHandle.filteredKeys);
  }
  init();
})();
</script>

## Yield distribution by cultivar

This chart responds to the filters above, including recommendation rank.
For example, with only rank 1 selected, it shows the yield distribution of
each location's single top-recommended combination per cultivar, rather
than every simulated planting-date treatment.

```{r boxplot}
# Individual jittered points (box outline/fill made transparent) rather than
# a plain aggregated box: crosstalk's filter response subsets each trace's
# underlying data and redraws, which should recompute box statistics too,
# but a shift in quartiles can be subtle to notice at a glance - showing
# every point makes "the filter actually changed what's plotted" obvious
# (fewer/more dots), rather than relying on a possibly-subtle box-shape shift.
plotly::plot_ly(
  sd, x = ~Cultivar, y = ~HWAH, color = ~Cultivar, type = "box",
  boxpoints = "all", jitter = 0.4, pointpos = 0,
  fillcolor = "rgba(0,0,0,0)", line = list(color = "rgba(0,0,0,0)"),
  showlegend = FALSE) %>%
  plotly::layout(
    xaxis = list(title = "Cultivar"),
    yaxis = list(title = "Yield (kg/ha)"))
```

## Recommendations table

For each location, the model evaluated 8 candidate planting dates
("treatments"), numbered 1–8 in chronological order. **Treatment 1**
corresponds to that location's estimated rainy-season onset — the first day,
on or after the 1st of the target planting month, on which at least 20 mm of
rain falls within a 3-day window and no 10-day dry spell (10 or more
consecutive days receiving less than 1 mm of rain) occurs within the
following 20 days. Treatments 2 through 8 are subsequent candidate planting
dates, spaced one week apart.

```{r table}
hide_idx <- which(names(dat) == "row_key") - 1

DT::datatable(
  sd,
  rownames = FALSE,
  colnames = c(
    "Zone" = "zone", "Rank" = "Combination_Rank", "Treatment" = "TRNO",
    "Latitude" = "XLAT", "Longitude" = "LONG", "Cultivar" = "Cultivar",
    "Planting date" = "PDAT", "Yield (kg/ha)" = "HWAH", "Biomass (kg/ha)" = "CWAM",
    "In crop mask" = "in_crop_mask"),
  extensions = "Buttons",
  options = list(
    dom = "Blfrtip",
    buttons = c("csv", "excel"),
    pageLength = 15,
    columnDefs = list(list(visible = FALSE, targets = hide_idx))))
```
