---
title: "`r params$crop_name` — DSSAT results dashboard"
output:
  html_document:
    self_contained: true
    theme:
      bootswatch: flatly
      version: 5
    mathjax: null
params:
  data_rds: ""
  crop_name: ""
  country_name: ""
  has_crop_mask: FALSE
---

<style>
/* rmarkdown's default template caps content at 940px regardless of theme -
   override it so the filters/maps/table can use more of a wide monitor. */
.main-container { max-width: 1600px; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressPackageStartupMessages({
  library(crosstalk)
  library(leaflet)
  library(plotly)
  library(DT)
  library(dplyr)
  library(ggplot2)
  library(htmltools)
})
```

```{r map-helpers, include=FALSE}
# Builds one independent leaflet widget bound once to `sd`. Deliberately
# exactly one crosstalk-bound addCircleMarkers layer per widget INSTANCE -
# leaflet's crosstalk filter handling (htmlwidgets/assets/leaflet.js) tracks
# marker visibility per (crosstalk group, row key) assuming a single layer
# per key; binding multiple layers to the same group with the same row_key
# values within ONE widget (as an earlier version of this file did, for a
# switchable single-map coloring) breaks its internal bookkeeping and
# silently stops filtering from working at all. Calling this function
# multiple times - once per map below - produces separate widgets, each
# with its own independent JS layer registry, so there's no collision.
#
# Marker colors use a FIXED scale (the full crop's data range for that
# metric) on every map, kept constant across filter changes so a given
# color always means the same thing no matter what's filtered - only each
# map's live text readout (built later, via the map-lookups/script blocks)
# reflects the current filter state, rather than re-coloring markers, since
# dynamically restyling already-rendered canvas markers would need custom
# JS reaching into leaflet's internal layer registry that can't be verified
# without a live browser.
build_dssat_map <- function(sd, element_id, color_formula, pal_fn, pal_values,
                             legend_title, label_formula, outline_parts,
                             leg_lab_format = NULL) {
  m <- leaflet::leaflet(
      sd, elementId = element_id, height = 480,
      options = leaflet::leafletOptions(preferCanvas = TRUE)) %>%
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) %>%
    leaflet::addCircleMarkers(
      lng = ~LONG, lat = ~XLAT, radius = 5, stroke = FALSE, fillOpacity = 0.8,
      color = color_formula, label = label_formula)

  m <- if (is.null(leg_lab_format)) {
    leaflet::addLegend(m, "bottomright", pal = pal_fn, values = pal_values, title = legend_title)
  } else {
    leaflet::addLegend(m, "bottomright", pal = pal_fn, values = pal_values,
                        title = legend_title, labFormat = leg_lab_format)
  }

  for (part in outline_parts) {
    m <- leaflet::addPolylines(
      m, lng = part$long, lat = part$lat, color = "black", weight = 1.5, opacity = 0.7)
  }
  m
}

# One small bordered column: a label, a map widget, and its live readout -
# keeps the three maps visually grouped together for the "check rank 1, see
# all three at once" workflow, with each map's label sitting right with it
# rather than as a page-level heading.
map_col <- function(label, map_widget, readout_id, readout_label) {
  htmltools::div(
    htmltools::tags$h5(label),
    map_widget,
    htmltools::div(
      style = "margin-top: 4px; font-weight: bold; font-size: 0.9em;",
      paste0(readout_label, ": "), htmltools::tags$span(id = readout_id, "computing…"))
  )
}
```

```{r load-data}
dat <- readRDS(params$data_rds)

cultivar_order <- intersect(c("Short", "Medium", "Long", "Longer"), unique(as.character(dat$Cultivar)))
dat$Cultivar <- factor(dat$Cultivar, levels = cultivar_order)
dat$PDAT <- as.Date(dat$PDAT)

display_cols <- c(
  "zone", "Combination_Rank", "TRNO", "XLAT", "LONG", "Cultivar", "PDAT",
  "HWAH", "CWAM", "in_crop_mask", "row_key")
dat <- dat[, display_cols]

sd <- crosstalk::SharedData$new(dat, key = ~row_key, group = "dssat")
```

# `r params$crop_name` — DSSAT results dashboard

This dashboard summarizes DSSAT crop model simulations for
**`r params$crop_name`**, driven by seasonal forecast weather data (not
historical/observed weather). At each pixel, the model was run for 8
planting-date treatments — spaced 7 days apart, starting from that pixel's
forecasted rainy-season onset — crossed with `r length(cultivar_order)`
cultivar maturity classes (`r paste(cultivar_order, collapse = ", ")`), for up
to `r 8 * length(cultivar_order)` simulated planting-date/cultivar
combinations per pixel. Open this file directly in a browser (no server or R
install required); it works fully offline except for the optional background
map imagery. Other crops in this usecase have their own separate dashboard
file, one per crop, in their own result folder.

`r if (!isTRUE(params$has_crop_mask)) paste0("**Note:** no crop-mask layer is available yet for ", params$crop_name, " — the \"Crop mask\" filter below has no effect (all pixels are always shown).")`

## Filters

```{r filters}
crosstalk::bscols(
  widths = c(6, 6),
  crosstalk::filter_checkbox("mask_filter", "Crop mask (checked = pixel is inside the mask)", sd, ~in_crop_mask),
  crosstalk::filter_checkbox("cultivar_filter", "Cultivar", sd, ~Cultivar)
)
```

<script>
// Crop mask defaults to checked on load. crosstalk::filter_checkbox has no
// default-checked parameter, so this simulates a user click: waiting for
// `load` (not `DOMContentLoaded`) guarantees htmlwidgets has already
// attached its delegated change listener before the simulated click fires.
window.addEventListener('load', function() {
  var input = document.querySelector('#mask_filter input[value="TRUE"]');
  if (input && !input.checked) {
    input.checked = true;
    input.dispatchEvent(new Event('change', { bubbles: true }));
  }
});
</script>

**Recommendation rank**: 1 = best yield at that pixel; higher numbers are
progressively worse alternatives. Check **1** to see only each pixel's top
recommendation; leave unchecked to see every combination.

```{r rank-filter}
crosstalk::filter_checkbox("rank_filter", "Recommendation rank", sd, ~Combination_Rank, inline = TRUE)
```

## Maps

```{r maps}
label_formula <- ~sprintf("%s | %s | %.0f kg/ha", Cultivar, format(PDAT, "%Y-%m-%d"), HWAH)

outline_raw <- tryCatch(
  ggplot2::map_data("world", region = params$country_name),
  error = function(e) NULL)
outline_parts <- if (!is.null(outline_raw) && nrow(outline_raw) > 0) {
  split(outline_raw, outline_raw$group)
} else {
  list()
}

pal_yield <- leaflet::colorNumeric("YlOrRd", domain = dat$HWAH, na.color = "transparent")
pal_cultivar <- leaflet::colorFactor("Dark2", domain = cultivar_order, na.color = "transparent")
pal_pdat <- leaflet::colorNumeric("RdYlBu", domain = as.numeric(dat$PDAT), na.color = "transparent", reverse = TRUE)
pdat_leg_labFormat <- function(type, cuts, p) format(as.Date(cuts, origin = "1970-01-01"), "%Y-%m-%d")

yield_map <- build_dssat_map(
  sd, "dssat_map_yield", ~pal_yield(HWAH), pal_yield, dat$HWAH,
  "Yield (kg/ha)<br>(fixed color scale)", label_formula, outline_parts)

cultivar_map <- build_dssat_map(
  sd, "dssat_map_cultivar", ~pal_cultivar(Cultivar), pal_cultivar, cultivar_order,
  "Cultivar<br>(fixed color scale)", label_formula, outline_parts)

pdat_map <- build_dssat_map(
  sd, "dssat_map_pdat", ~pal_pdat(as.numeric(PDAT)), pal_pdat, as.numeric(dat$PDAT),
  "Planting date<br>(fixed color scale)", label_formula, outline_parts,
  leg_lab_format = pdat_leg_labFormat)

crosstalk::bscols(
  widths = c(4, 4, 4),
  map_col("Estimated Yields", yield_map, "yield-range-text", "Visible yield range"),
  map_col("Recommended Cultivar", cultivar_map, "cultivar-counts-text", "Visible cultivar counts"),
  map_col("Recommended Planting Date", pdat_map, "pdat-range-text", "Visible planting-date range")
)
```

```{r map-lookups, include=FALSE}
# key -> value lookups consumed by the dynamic-legend script below. Kept out
# of the visible chunks (include=FALSE) since they're only for the JS, not
# for display.
yield_by_key_json <- jsonlite::toJSON(setNames(as.list(dat$HWAH), dat$row_key), auto_unbox = TRUE)
cultivar_by_key_json <- jsonlite::toJSON(setNames(as.list(as.character(dat$Cultivar)), dat$row_key), auto_unbox = TRUE)
# Plain ISO YYYY-MM-DD strings, not JS Date objects - they sort correctly
# with ordinary string min/max, so no date parsing is needed client-side.
pdat_by_key_json <- jsonlite::toJSON(setNames(as.list(format(dat$PDAT, "%Y-%m-%d")), dat$row_key), auto_unbox = TRUE)
cultivar_levels_json <- jsonlite::toJSON(cultivar_order)
```

<script>
window.dssatYieldByKey = `r yield_by_key_json`;
window.dssatCultivarByKey = `r cultivar_by_key_json`;
window.dssatPdatByKey = `r pdat_by_key_json`;
window.dssatCultivarLevels = `r cultivar_levels_json`;

(function() {
  function fmt(x) { return Math.round(x).toLocaleString(); }

  function valuesFor(lookup, keys) {
    if (keys === null || typeof keys === "undefined") {
      return Object.values(lookup);
    }
    return keys
      .map(function(k) { return lookup[k]; })
      .filter(function(v) { return typeof v !== "undefined"; });
  }

  function updateYield(keys) {
    var el = document.getElementById("yield-range-text");
    if (!el) return;
    var vals = valuesFor(window.dssatYieldByKey, keys).filter(function(v) { return typeof v === "number" && !isNaN(v); });
    if (vals.length === 0) { el.textContent = "no pixels match the current filters"; return; }
    var lo = Math.min.apply(null, vals);
    var hi = Math.max.apply(null, vals);
    el.textContent = fmt(lo) + " – " + fmt(hi) + " kg/ha (" + vals.length + " points)";
  }

  function updatePdat(keys) {
    var el = document.getElementById("pdat-range-text");
    if (!el) return;
    var vals = valuesFor(window.dssatPdatByKey, keys).filter(function(v) { return typeof v === "string"; });
    if (vals.length === 0) { el.textContent = "no pixels match the current filters"; return; }
    vals.sort();
    el.textContent = vals[0] + " – " + vals[vals.length - 1] + " (" + vals.length + " points)";
  }

  function updateCultivar(keys) {
    var el = document.getElementById("cultivar-counts-text");
    if (!el) return;
    var vals = valuesFor(window.dssatCultivarByKey, keys).filter(function(v) { return typeof v === "string"; });
    var counts = {};
    window.dssatCultivarLevels.forEach(function(lvl) { counts[lvl] = 0; });
    vals.forEach(function(v) { if (counts.hasOwnProperty(v)) counts[v] += 1; });
    el.textContent = window.dssatCultivarLevels
      .map(function(lvl) { return lvl + ": " + counts[lvl]; })
      .join(" · ");
  }

  function updateAll(keys) {
    updateYield(keys);
    updatePdat(keys);
    updateCultivar(keys);
  }

  // Mirrors the exact crosstalk.FilterHandle pattern leaflet's own bundled
  // JS uses internally (new crosstalk.FilterHandle(group), .on("change", ...),
  // reading .filteredKeys for the initial state) - reused here directly
  // rather than reverse-engineered from scratch, and confirmed (by reading
  // crosstalk.js's EventEmitter-based implementation) that a single
  // FilterHandle can safely drive multiple readouts from one subscription.
  function init() {
    if (typeof crosstalk === "undefined") { setTimeout(init, 200); return; }
    var filterHandle = new crosstalk.FilterHandle("dssat");
    filterHandle.on("change", function(e) { updateAll(e.value); });
    updateAll(filterHandle.filteredKeys);
  }
  init();
})();
</script>

## Yield distribution by cultivar

This plot responds to the same filters above, including recommendation rank —
so if only rank 1 is checked, it shows the yield distribution of just each
pixel's top choice per cultivar, not every planting-date treatment.

```{r boxplot}
# Individual jittered points (box outline/fill made transparent) rather than
# a plain aggregated box: crosstalk's filter response subsets each trace's
# underlying data and redraws, which should recompute box statistics too,
# but a shift in quartiles can be subtle to notice at a glance - showing
# every point makes "the filter actually changed what's plotted" obvious
# (fewer/more dots), rather than relying on a possibly-subtle box-shape shift.
plotly::plot_ly(
  sd, x = ~Cultivar, y = ~HWAH, color = ~Cultivar, type = "box",
  boxpoints = "all", jitter = 0.4, pointpos = 0,
  fillcolor = "rgba(0,0,0,0)", line = list(color = "rgba(0,0,0,0)"),
  showlegend = FALSE) %>%
  plotly::layout(
    xaxis = list(title = "Cultivar"),
    yaxis = list(title = "Yield (kg/ha)"))
```

## Recommendations table

```{r table}
hide_idx <- which(names(dat) == "row_key") - 1

DT::datatable(
  sd,
  rownames = FALSE,
  colnames = c(
    "Zone" = "zone", "Rank" = "Combination_Rank", "Treatment" = "TRNO",
    "Latitude" = "XLAT", "Longitude" = "LONG", "Cultivar" = "Cultivar",
    "Planting date" = "PDAT", "Yield (kg/ha)" = "HWAH", "Biomass (kg/ha)" = "CWAM",
    "In crop mask" = "in_crop_mask"),
  extensions = "Buttons",
  options = list(
    dom = "Blfrtip",
    buttons = c("csv", "excel"),
    pageLength = 15,
    columnDefs = list(list(visible = FALSE, targets = hide_idx))))
```
